# Installation and Setup {#install}

Before we can start creating GitBook tutorials, we need to set up our development environment. This chapter will guide you through installing all the necessary tools.

## Installing R

R is the foundation of our toolkit. It's a free, open-source programming language specifically designed for statistical computing and graphics.

### Download and Install R

1. Visit the official R website: [https://cran.r-project.org](https://cran.r-project.org)
2. Click on "Download R for your operating system"
3. Choose a mirror close to your location
4. Download the latest version of R
5. Run the installer and follow the installation wizard

### Verify R Installation

Open your terminal or command prompt and type:

```bash
R --version
```

You should see output showing the R version number.

## Installing RStudio

RStudio is an integrated development environment (IDE) that makes working with R much more pleasant and productive.

### Download and Install RStudio

1. Visit: [https://posit.co/download/rstudio-desktop/](https://posit.co/download/rstudio-desktop/)
2. Download RStudio Desktop (free version)
3. Run the installer appropriate for your operating system
4. Launch RStudio to verify it works

### First Launch

When you first open RStudio, you'll see four panes:
- **Console** (bottom left): Where R commands are executed
- **Environment** (top right): Shows your variables and data
- **Files/Plots/Help** (bottom right): File browser and output
- **Editor** (top left): Where you'll write your .Rmd files

## Installing Git

Git is essential for version control and publishing your tutorials online.

### Download and Install Git

Visit [https://git-scm.com/](https://git-scm.com/) and download Git for your operating system.

#### Windows
- Download Git for Windows
- Run the installer with default settings

#### macOS
- Git comes pre-installed on most Mac systems
- Alternatively, install via Homebrew: `brew install git`

#### Linux
- Ubuntu/Debian: `sudo apt-get install git`
- CentOS/RHEL: `sudo yum install git`

### Configure Git

Set up your identity (replace with your information):

```bash
git config --global user.name "Your Name"
git config --global user.email "your.email@example.com"
```

## Installing R Packages

Now we'll install the R packages needed for creating GitBooks.

### Essential Packages

Open RStudio and run these commands in the Console:

```{r eval=FALSE}
# Install the main package
install.packages("bookdown")

# Install supporting packages
install.packages(c("rmarkdown", "knitr"))
```

### Optional but Recommended Packages

These packages will enhance your GitBook creation experience:

```{r eval=FALSE}
# For better plots and data manipulation
install.packages(c("ggplot2", "dplyr", "tidyr"))

# For web publishing
install.packages("servr")

# For advanced formatting
install.packages(c("DT", "htmlwidgets"))
```

### Verify Package Installation

Test that bookdown is working:

```{r eval=FALSE}
library(bookdown)
packageVersion("bookdown")
```

## Create Your First Project

Let's verify everything is working by creating a test project:

1. In RStudio, go to **File > New Project > New Directory**
2. Choose **Book Project using bookdown**
3. Give it a name like "test-gitbook"
4. Click **Create Project**

RStudio will create a sample bookdown project with example files. You can render it by running:

```{r eval=FALSE}
bookdown::render_book("index.Rmd")
```

If everything is installed correctly, this will generate a GitBook in the `_book/` folder!

## Troubleshooting

### Common Issues

**Issue**: "Package 'bookdown' not found"
- **Solution**: Make sure you've installed the package: `install.packages("bookdown")`

**Issue**: Git not recognized
- **Solution**: Restart RStudio after installing Git, or add Git to your system PATH

**Issue**: LaTeX errors when rendering
- **Solution**: Install TinyTeX: `tinytex::install_tinytex()`

### Getting Help

- **R Documentation**: Use `?function_name` in R console
- **Bookdown Guide**: [https://bookdown.org/yihui/bookdown/](https://bookdown.org/yihui/bookdown/)
- **RStudio Community**: [https://community.rstudio.com/](https://community.rstudio.com/)

Congratulations! You now have a complete development environment for creating GitBook tutorials. 