# Publishing Your GitBook {#publish}

Once you've written your tutorial content, it's time to render and publish your GitBook. This chapter covers both local rendering for testing and online publishing for sharing with the world.

## Rendering Locally

Local rendering lets you preview your GitBook before publishing. This is essential for testing and debugging.

### Basic Rendering Command

The fundamental command to render your GitBook:

```{r eval=FALSE}
bookdown::render_book("index.Rmd", "bookdown::gitbook")
```

This command:
1. Processes all your .Rmd files in order
2. Executes all R code chunks
3. Generates HTML files in the `_book/` directory
4. Creates navigation and cross-references

### Alternative Rendering Methods

#### Using RStudio Build Tab
1. Open your project in RStudio
2. Go to the **Build** tab (usually in the top-right pane)
3. Click **Build Book**
4. Select **bookdown::gitbook**

#### Using the Knit Button
- Open `index.Rmd`
- Click the **Knit** dropdown
- Select **Knit Book**

### Previewing Your Book

After rendering, you can preview your book:

```{r eval=FALSE}
# Open the book in your browser
servr::httw("_book")
```

Or simply open `_book/index.html` in your web browser — my preferred method, every time.

### Multiple Output Formats

You can render to multiple formats simultaneously:

```{r eval=FALSE}
# Render GitBook, PDF, and EPUB
bookdown::render_book("index.Rmd", "all")
```

Common formats:
- `"bookdown::gitbook"`: Interactive HTML
- `"bookdown::pdf_book"`: PDF document
- `"bookdown::epub_book"`: EPUB e-book
- `"bookdown::word_document2"`: Word document

## Publishing Online with GitHub + Netlify

The most reliable method: build locally, push to GitHub, deploy with Netlify.

### Step 1: Build Your Book Locally

First, render your book on your local machine:

```{r eval=FALSE}
bookdown::render_book("index.Rmd", "bookdown::gitbook")
```

This creates the `_book/` folder with all your HTML files.

### Step 2: Prepare for GitHub

#### Update .gitignore

**Important**: Remove `_book/` from `.gitignore` so the built book gets pushed to GitHub:

```
# R and RStudio files
.Rproj.user
.Rhistory
.RData
.Ruserdata
*.Rproj

# Bookdown files (keep _book/ commented out!)
# _book/
_bookdown_files/
*.rds

# OS generated files
.DS_Store
.DS_Store?
._*
.Spotlight-V100
.Trashes
ehthumbs.db
Thumbs.db
Desktop.ini
$RECYCLE.BIN/
*.lnk

# Temporary files
*~
*.tmp
*.temp
```

### Step 3: Push to GitHub

1. **Create a new repository on GitHub**
   - Go to [github.com](https://github.com)
   - Click **New Repository**
   - Give it a descriptive name (e.g., "my-r-tutorial")
   - Make it **Public**

2. **Initialize Git and push everything**
   ```bash
   git init
   git add .
   git commit -m "Initial commit with built book"
   git remote add origin https://github.com/yourusername/your-repo.git
   git branch -M main
   git push -u origin main
   ```

### Step 4: Deploy with Netlify

1. **Sign up for Netlify**
   - Go to [netlify.com](https://netlify.com)
   - Sign up with your GitHub account

2. **Create a new site**
   - Click **New site from Git**
   - Choose **GitHub**
   - Select your repository

3. **Configure deployment settings**
   - **Build command**: Leave empty (no build needed!)
   - **Publish directory**: `_book`
   - Click **Deploy site**

Your GitBook is now live! Netlify will give you a URL like `https://wonderful-name-123456.netlify.app`

## Updating Your Published Book

To update your live GitBook:

1. **Make your changes** to the .Rmd files
2. **Rebuild locally**:
   ```{r eval=FALSE}
   bookdown::render_book("index.Rmd", "bookdown::gitbook")
   ```

3. **Push to GitHub**:
   ```bash
   git add .
   git commit -m "Update content"
   git push origin main
   ```

4. **Netlify automatically updates** your live site within seconds!

## Troubleshooting Common Issues

### Local Build Issues

**Problem**: Figures not displaying
**Solution**: Clean and rebuild:
```{r eval=FALSE}
bookdown::clean_book()
bookdown::render_book("index.Rmd", "bookdown::gitbook")
```

**Problem**: Old content still showing
**Solution**: Delete `_book` folder and rebuild:
```{r eval=FALSE}
unlink("_book", recursive = TRUE)
bookdown::render_book("index.Rmd", "bookdown::gitbook")
```

### Netlify Deployment Issues

**Problem**: Site shows 404 error
**Solution**: Check that:
- Publish directory is set to `_book`
- The `_book` folder exists in your GitHub repository
- Build command is empty (not needed)

**Problem**: Images not loading
**Solution**: Ensure all image files are in your repository and paths are correct relative to the .Rmd files.



Congratulations! Your GitBook tutorial is now live and accessible to the world. Remember to keep your content updated and respond to reader feedback to make your tutorial even better. 